\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-11-09 doc/flexiblesusy-paper.tex
fix typographical errors in the paper

%macros
\newcommand{\sarah}{SARAH\@\xspace}
\newcommand{\fs}{FlexibleSUSY\@\xspace}
\newcommand{\mathematica}{Mathematica\xspace}
\newcommand{\ESSM}{E$_6$SSM\@\xspace}
\newcommand{\code}[1]{\lstinline|#1|}  % inline source code
\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
\newcommand{\DRbar}{\textoverline{DR}\xspace}
\newcommand{\MSbar}{\textoverline{MS}\xspace}
\newcommand{\unit}[1]{\,\text{#1}}      % units
\newcommand{\userinput}{\text{input}}
\newcommand{\pole}{\text{pole}}
\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\unity}{\mathbf{1}}
\newcommand{\figref}[1]{\figurename~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\tabref}[1]{\tablename~\ref{#1}}
\newcommand{\ptitle}[1]{\emph{#1}}
\renewcommand{\ptitle}[1]{}
\newcommand{\Zv}{\mathbf{\backslash}\mkern-11.0mu{Z}}
\makeatletter
\lstnewenvironment{numlstlisting}[1][]{%
  \lstset{%
    #1,
    numbers=left,
    firstnumber=auto,
    numberstyle=\tiny\sffamily}%
  \csname\@lst @SetFirstNumber\endcsname
}{%
  \csname \@lst @SaveFirstNumber\endcsname
}
\makeatother
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\re}{Re}
\DeclareMathOperator{\im}{Im}
\def\at{\alpha_t}
\def\ab{\alpha_b}
\def\as{\alpha_s}
\def\atau{\alpha_{\tau}}
\def\oat{O(\at)}
\def\oab{O(\ab)}
\def\oatau{O(\atau)}
\def\oatab{O(\at\ab)}
\def\oatas{O(\at\as)}
\def\oabas{O(\ab\as)}
\def\oatababq{O(\at\ab + \ab^2)}
\def\oatqatababq{O(\at^2 + \at\ab + \ab^2)}
\def\oatasatq{O(\at\as + \at^2)}
\def\oatasabas{O(\at\as +\ab\as)}
\def\oatasabasatq{O(\at\as + \at^2 +\ab\as)}
\def\oatq{O(\at^2)}
\def\oabq{O(\ab^2)}
\def\oatauq{O(\atau^2)}
\def\oabatau{O(\ab \atau)}
\def\oatplusabsq{O((\at+\ab)^2)}
\def\oas{O(\as)}
\def\oatauqatab{O(\atau^2 +\ab \atau )}

@@

To check the correctness of \fs's generated spectrum generators
extensive unit testing against Softsusy's MSSM and NMSSM
implementations (both $Z_3$-invariant and $Z_3$-violating variants)
has been carried out.  These unit tests systematically compare all
tree-level mass matrices, EWSB equations, one- and two-loop
$\beta$-functions, one- and two-loop self-energies and one- and
two-loop tadpoles numerically for the CMSSM and NMSSM parameter points
given in \tabref{tab:unit-test-parameter-points}.
%
\begin{table}[tbh]
  \centering
  \begin{tabularx}{\textwidth}{lX}
    \toprule
    SUSY Model & Tested parameter points\\
    \midrule CMSSM & $m_0 = 125\unit{GeV}$, $M_{1/2} = 500\unit{GeV}$,
    $\tan\beta = 10$, $A_0 = 0$, $\sign\mu = \pm 1$
    \\
    $Z_3$-NMSSM & $m_0 = \{250,300\}\unit{GeV}$, $M_{1/2} =
    200\unit{GeV}$, $\tan\beta = 10$, $A_0 = -500\unit{GeV}$,
    $\sign\mu_\text{eff} = +1$, $\lambda = 0.1$
    \\
    $\Zv_3$-NMSSM & $m_0 = 540\unit{GeV}$, $M_{1/2} = 200\unit{GeV}$,
    $\tan\beta = 10$, $A_0 = -350\unit{GeV}$, $\sign\mu_\text{eff} =
    \pm 1$, $\lambda = \kappa = 0.1$, $s = 1\unit{TeV}$, $\mu' =
    290\unit{GeV}$, $m_S^{\prime 2} = 400\unit{GeV}$, $\xi_F =
    300\unit{GeV}$
    \\
    \bottomrule
  \end{tabularx}
  \caption{CMSSM and NMSSM parameter points used for the unit tests
    against Softsusy.
    We follow the notation of the NMSSM model parameters used in
    \cite{Ellwanger:2009dp,Allanach:2013kza}.}
  \label{tab:unit-test-parameter-points}
\end{table}
%
All tested expressions were found to agree within double machine
precision.\footnote{Due to the systematic and detailed tests several
  bugs in Softsusy, SARAH and \fs could be identified and corrected.}
Furthermore, the output of the iterative procedures which solve the
one- and two-loop corrected tadpole equations
\eqref{eq:one-loop-ewsb-eq} to find the minimum of the effective Higgs
potential were compared numerically for these parameter points and
found to agree within machine precision as well.  Finally, the overall
pole mass spectrum and mixing after the full fixed-point iteration has
finished has been compared, and was found to agree at the sub-permille
level.  The origin of the sub-permille level difference between \fs
and Softsusy is the different determination of the weak mixing angle
$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the SUSY model in the
\DRbar\ scheme: \fs calculates
$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from $M_W$ and $M_Z$, as
described in \secref{sec:calculation-of-gauge-couplings}, while
Softsusy determines $\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from
the muon decay constant $G_\mu$.  The approach used in Softsusy
results in more precise \DRbar\ gauge couplings at the $M_Z$ scale,
because the muon decay constant is known with a higher accuracy than
the $W$-boson mass.  Furthermore, in Softsusy $3.5.0$ some three-loop
$\beta$-functions and two-loop threshold corrections can be enabled in
the MSSM to increase the accuracy of the RG running and the
determination of the \DRbar\ gauge and Yukawa couplings at $M_Z$
\cite{Allanach:2014nba}.  These corrections are not implemented in \fs
so far.  The complete set of unit tests is shipped with \fs and can be
found in the \code{test/} directory.  The tests can be run with the
command \code{make execute-tests}.  All unit tests are carried out
nightly in order to continuously check the correctness of the meta
code and the generated spectrum generators for the shipped models.
The nightly test results can be found at
\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}

\section{Examples of C++ code adaptation}
\label{sec:examples of c++ code adaptation}

\subsection{Stacking models in a tower of effective theories}
\label{sec:tower construction}

@@

\subsection{Integrating custom-built C++ components}
\label{sec:integrating-custom-built}

@@

>>>>

%macros
\newcommand{\sarah}{SARAH\@\xspace}
\newcommand{\fs}{FlexibleSUSY\@\xspace}
\newcommand{\mathematica}{Mathematica\xspace}
\newcommand{\ESSM}{E$_6$SSM\@\xspace}
\newcommand{\code}[1]{\lstinline|#1|}  % inline source code
\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
\newcommand{\DRbar}{\textoverline{DR}\xspace}
\newcommand{\MSbar}{\textoverline{MS}\xspace}
\newcommand{\unit}[1]{\,\text{#1}}      % units
\newcommand{\userinput}{\text{input}}
\newcommand{\pole}{\text{pole}}
\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\unity}{\mathbf{1}}
\newcommand{\figref}[1]{\figurename~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\tabref}[1]{\tablename~\ref{#1}}
\newcommand{\ptitle}[1]{\emph{#1}}
\renewcommand{\ptitle}[1]{}
\makeatletter
\lstnewenvironment{numlstlisting}[1][]{%
  \lstset{%
    #1,
    numbers=left,
    firstnumber=auto,
    numberstyle=\tiny\sffamily}%
  \csname\@lst @SetFirstNumber\endcsname
}{%
  \csname \@lst @SaveFirstNumber\endcsname
}
\makeatother
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\re}{Re}
\DeclareMathOperator{\im}{Im}
\def\at{\alpha_t}
\def\ab{\alpha_b}
\def\as{\alpha_s}
\def\atau{\alpha_{\tau}}
\def\oat{O(\at)}
\def\oab{O(\ab)}
\def\oatau{O(\atau)}
\def\oatab{O(\at\ab)}
\def\oatas{O(\at\as)}
\def\oabas{O(\ab\as)}
\def\oatababq{O(\at\ab + \ab^2)}
\def\oatqatababq{O(\at^2 + \at\ab + \ab^2)}
\def\oatasatq{O(\at\as + \at^2)}
\def\oatasabas{O(\at\as +\ab\as)}
\def\oatasabasatq{O(\at\as + \at^2 +\ab\as)}
\def\oatq{O(\at^2)}
\def\oabq{O(\ab^2)}
\def\oatauq{O(\atau^2)}
\def\oabatau{O(\ab \atau)}
\def\oatplusabsq{O((\at+\ab)^2)}
\def\oas{O(\as)}
\def\oatauqatab{O(\atau^2 +\ab \atau )}

@@

\subsubsection{Stacking models in a tower of effective theories}
\label{sec:tower construction}

@@

\subsubsection{Integrating custom-built C++ components}
\label{sec:integrating-custom-built}

@@

To check the correctness of \fs's generated spectrum generators
extensive unit testing against Softsusy's MSSM and NMSSM
implementations (both $Z_3$-invariant and $Z_3$-violating variants)
has been carried out.  These unit tests systematically compare all
tree-level mass matrices, EWSB equations, one- and two-loop
$\beta$-functions, one- and two-loop self-energies and one- and
two-loop tadpoles numerically for various MSSM and NMSSM parameter
points.  All tested expressions were found to agree within double
machine precision.\footnote{Due to the systematic and detailed tests
  several bugs in Softsusy, SARAH and \fs could be identified and
  corrected.}  Furthermore, the overall pole mass spectrum after the
full fixed-point iteration has finished has been compared, and was
found to agree at the sub-permille level.  The origin of the
sub-permille level difference between \fs and Softsusy is the
different determination of the weak mixing angle
$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the SUSY model in the
\DRbar\ scheme: \fs calculates
$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from $M_W$ and $M_Z$, as
described in \secref{sec:calculation-of-gauge-couplings}, while
Softsusy determines $\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from
the muon decay constant $G_\mu$.  The approach used in Softsusy
results in more precise \DRbar\ gauge couplings at the $M_Z$ scale,
because the muon decay constant is known with a higher accuracy than
the $W$-boson mass.  Furthermore, in Softsusy $3.5.0$ some three-loop
$\beta$-functions and two-loop threshold corrections can be enabled in
the MSSM to increase the accuracy of the RG running and the
determination of the \DRbar\ gauge and Yukawa couplings at $M_Z$
\cite{Allanach:2014nba}.  These corrections are not implemented in \fs
so far.  The complete set of unit tests is shipped with \fs and can be
found in the \code{test/} directory.  The tests can be run with the
command \code{make execute-tests}.  All unit tests are carried out
nightly in order to continuously check the correctness of the meta
code and the generated spectrum generators for the shipped models.
The nightly test results can be found at
\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}
